---
title: "A3: SPL Library Checkouts"
output: html_document
---
### Introduction

An introduction of the data and a description of the trends/books/items you are choosing to analyze (and why!)

### Summary Information

Write a summary paragraph of findings that includes the 5 values calculated from your summary information R script

For this assignment, I analyzed all PHYSICAL checkouts from the library through 2022. In my analysis
I discovered that the most popular author of 2022 was Emily St. John Mandel, with the most popular book
being of the year _Sea of Tranquility_ written by her. I also learned, the most popular month for checking out books was January, most likely due to New Years resolutions towards reading. Lastly, I discovered that behind
book checkouts, video and sound discs were the most commonly checked out material type.

### The Dataset

- Who collected/published the data?  

The Seattle Public Library

- What are the parameters of the data (dates, number of checkouts, kinds of books, etc.)?

They refresh monthly, but keep track of way more
than just book checkouts. They also collect data
on items checked out online. It goes back to Apr2005

- How was the data collected or generated?  

This data was collected using a monthly count
of Seattle Public Library checkouts of both
physical and electronic items

- Why was the data collected?

They claim to collect this data to: improve
public understanding, generate economic 
opportunity, empower city employees, and
encourage development.

- What, if any, ethical questions do you need to consider when working with this data?  

I do not see any ethical concerns when it comes
to this data; however, if this data starts to 
include PERSONAL information of those who checkout
books, or if the data is used towards book banning
or de-funding the library, then I think it starts
to enter an unethical territory.

- What are possible limitations or problems with this data?   (at least 200 words)

One that comes to mind is we don't know who is
checking out what. For instance, one individual
may be using the library extremely frequently,
checking out many more books than the average 
individual, which will make statistics on checkouts 
more skewed to their interests.


### First Trends Over Time Chart
```{r, echo = FALSE}
library(dplyr)
library(ggplot2)
library(RColorBrewer)

books_per_month_df <- library_df %>% filter(MaterialType == "BOOK") %>% select(CheckoutMonth, Checkouts) %>% group_by(CheckoutMonth) %>% summarize(total_checkouts = sum(Checkouts))

ggplot(data = books_per_month_df) + 
      geom_line(aes(x=CheckoutMonth, y = total_checkouts)) +
      labs(title = "Books Checked Out Per Month", x = "Month", y = "Books Checked Out")


```

This first chart shows the trends of Books check out in the library per month. I chose this chart to show the change throughout the year for the most basic, fundamental use of the library. As can be seen by the chart, throughout the year, less and less people check out books from the library. 

### Second Trends Over Time Chart
```{r, echo = FALSE}
vDiscs_per_month_df <- library_df %>% filter(MaterialType == "VIDEODISC") %>% select(CheckoutMonth, Checkouts) %>% group_by(CheckoutMonth) %>% summarize(total_checkouts = sum(Checkouts))

ggplot(data = vDiscs_per_month_df) + 
  geom_line(aes(x=CheckoutMonth, y = total_checkouts)) +
  labs(title = "Video Discs Checked Out Per Month", x = "Month", y = "Video Discs Checked Out")

```

This second chart shows the trends of Video Discs checked out in the library per month. I chose this chart to see if the second highest checked out material, behind books, followed the same trends as books did. As can be seen by the chart, Video Discs follow the same trend as Books, meaning that overall less people are going to the library in person, and they are not switching mediums. 

### Your Choice
```{r, echo = FALSE}

material_ranking_df <- library_df %>% select(MaterialType, Checkouts) %>% group_by(MaterialType) %>% summarize(total_checkouts = sum(Checkouts)) %>% arrange(-total_checkouts) %>% slice_head(n=3)

ggplot(data = material_ranking_df) + 
  geom_col(aes(x = MaterialType, y = total_checkouts, fill = MaterialType)) +
  scale_color_brewer(palette = "Set2") +
  labs(title = "Material Total Checkout Comparisons", x = "Material Type", y = "Checkouts", fill = "MaterialType")

```

This last chart shows the top 3 material types checked out from the library. I chose this chart to give context to my previous charts, and why I chose to analyze book and video disc trends. As can be seen by the chart, books and video discs are the two most popular material types checked out from the library. 

